Parallel algorithm for solving multicriterial optimization problems using elements of machine learning

### 2025-06-17 2025/RSCD_MCO/RSD-2025.tex
minor correction

\begin{figure}
\center
\includegraphics[width=0.6\textwidth]{fig2.png}
\caption{Example of Pareto set evaluation in solving two scalar optimization problems (values in the domain of criteria values).} \label{fig2}
\end{figure}

@@

As a criterion for the quality of the found Pareto set, the indicator HV \cite{Evtushenko2014,pymoo,AGP_ML} was applied. Higher value of the indicator corresponds to better solution. To assess the performance of the algorithm, the number of trials carried out when solving the problem are given. However, when solving complex optimization problems, one trial can take a significant time. In such conditions, the reduction of the number of trials reduces the total computation time in general. In the presented experimental results, the number of trials spent on solving each problem and the HV indicators were averaged according to the number of solved problems.

@@

Fig. \ref{fig5} shows the dependence of the average value of the HV index on the average number of trials for the sequential algorithm. Several solution accuracy values $\varepsilon = \{0.1,0.05,0.01\}$ were fixed for plotting the graph for algorithm (\ref{eq:13}). 100 problems were solved for each accuracy value and the average results were plotted on a graph. The position of the resulting curve for the chosen algorithm to the left and above corresponds to the best behavior of the algorithm as a whole. During the experiments the parameter $\alpha$ from (15) varied, Fig. \ref{fig5} shows only the best results for each of the variants.

>>>>

\begin{figure}
\center
\includegraphics[width=0.6\textwidth]{fig2.png}
\caption{Example of Pareto set evaluation in solving two scalar optimization problems (trial points are presented in the criteria space).} \label{fig2}
\end{figure}

@@

As a criterion for the quality of the found Pareto set, the indicator HV was applied. Higher value of the indicator corresponds to better solution. 
HV index is a traditional quality metric when approximating a Pareto set \cite{Evtushenko2014,pymoo,AGP_ML}.
To assess the performance of the algorithm, the number of trials carried out when solving the problem are given. However, when solving complex optimization problems, one trial can take a significant time. In such conditions, the reduction of the number of trials reduces the total computation time in general. In the presented experimental results, the number of trials spent on solving each problem and the HV indicators were averaged according to the number of solved problems.

@@

Fig. \ref{fig5} shows the dependence of the average value of the HV index on the average number of trials for the sequential algorithm. The number of $\lambda$ values from (\ref{eq:04}) was equal to $\tau=50$ in all experiments. The $\lambda$ values were chosen to be uniformly distributed over the interval [0,1]. Several solution accuracy values $\varepsilon = \{0.1,0.05,0.01\}$ were fixed for plotting the graph for algorithm (\ref{eq:13}). 100 problems were solved for each accuracy value and the average results were plotted on a graph. The position of the resulting curve for the chosen algorithm to the left and above corresponds to the best behavior of the algorithm as a whole. During the experiments the parameter $\alpha$ from (15) varied within the range $[0.01,0.05]$ with a step of $0.01$, Fig. \ref{fig5} shows only the best results for each of the variants.
